---
title: "ProjectWork"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 **Importing the Data**
 
```{r}
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(plotly)
dat <- read_excel("ALL.xlsx")
dat <- select(dat, -Player2, -Player3)
```
Create Variables we need
```{r}
  dat <- mutate(dat, RBIAB = RBI / AB)
  dat$RBIAB <- round(dat$RBIAB, digits = 3)
  ##Runs batted in per at bat^
  dat <- mutate(dat, HRAB = HR / AB) 
  dat$HRAB <- round(dat$HRAB, digits = 3)
  ##Home Runs per at bat^
  dat <- mutate(dat, BABIP = (H - HR) / (AB - SO - HR + SF))
  dat$BABIP <- round(dat$BABIP, digits = 3)
  ##Batting Average on Balls in Play^
  dat$NP <- as.numeric(dat$NP) ##Cleaning the Data, changing column from character to Number
  dat <- mutate(dat, NPPA = NP / PA)
  dat$NPPA <- round(dat$NPPA, digits = 3)
  ##Number of pitches per plate appearance^
  dat <- mutate(dat, NPAB = NP / AB)
  dat$NPAB <- round(dat$NPAB, digits = 3)
  ##Number of Pitches per At bat per at bat^
  dat <- mutate(dat, SOAB = SO / AB)
  dat$SOAB <- round(dat$SOAB, digits = 3)
  ##Number of StrikeOuts per at bat^
```


Visualize the Data
```{r}
ggplot(dat, aes(x = AVG)) + geom_histogram(binwidth = .002,color="Black", fill="light blue") + xlab("Batting Average") + ggtitle("Histogram of Batting Average") + ylab("Count")

ggplot(dat, aes(x = OPS)) + geom_histogram(binwidth = .005,color="Black", fill="light green") + xlab("On-Base + Slugging Percentage") + ggtitle("Histogram of OPS") + ylab("Count")
str(dat)
correlationMatrix3 <- select(dat, SLG, RBIAB)
cor(correlationMatrix3)

ggplot(dat, aes(x = RBIAB, y = SLG)) + geom_point(color = 'coral2') + geom_smooth(method='lm',formula=y~x) + xlab("RBI per At bat") + ggtitle("Relationship between Slugging Percentage and RBI's per At Bat") + ylab("Slugging Percentage") + annotate("Text", .325, .55, label = "      R = .796 \n R^2 = .634")

correlationMatrix3 <- select(dat, RBI, SLG)
cor(correlationMatrix3)
```
Does getting more pitches in an at Bat increase the odds of hitting a homerun? *PROBABLY WONT USE THIS*
```{r}
ggplot(dat, aes(x = HRAB, y = NPPA)) + geom_point() + geom_smooth(method='lm',formula=y~x) ## R_SQUARED

```
Which position has the best batting average?
```{r}
dat2 <- dat
dat2%>%
  select(Pos, AVG) %>%
  mutate(AVG = ifelse(AVG == -9999, NA, AVG)) %>%
  group_by(Pos) %>%
  mutate(MEANAVG = mean(AVG, na.rm = T)) -> dat2
ggplot(dat2, aes(x = AVG)) + geom_histogram(binwidth = .002,color="Black", fill="goldenrod2") + xlab("Batting Average") + ggtitle("Histogram of Batting Average by Position") + ylab("Count") + facet_wrap(~Pos) + geom_vline(aes(xintercept = MEANAVG))

dat3 <- select(dat2, -AVG)
dat3 <- distinct(dat3)  #ROUND TO 3 digits
dat3$MEANAVG <- round(dat3$MEANAVG, digits = 3)

head(dat3, 11)
# WOW, This makes no sense, as the lowest batting average on the field is the DH and the Right fielder
```
Teams batting average?
```{r}
dat5 <- dat
dat5%>%
  select(Team, AVG) %>%
  mutate(AVG = ifelse(AVG == -9999, NA, AVG)) %>%
  group_by(Team) %>%
  mutate(MEANAVG = mean(AVG, na.rm = T)) -> dat5

dat6 <- select(dat5, -AVG)
dat6 <- distinct(dat6)
dat6$MEANAVG <- round(dat6$MEANAVG, digits = 3)

head(dat6, 10)
```

Following Albert Pujols through his career 
```{r}
Albert <- dat[grepl("Pujols,A", dat$Player),]

ggplot(Albert, aes(x = Year)) +
  geom_line(aes( y = AVG), color = "blue") +
  geom_line(aes( y = SLG), color = "red") + 
  geom_line(aes( y = HRAB), color = "purple") + 
  geom_line(aes( y = RBIAB), color = "Orange") +
  ylab("Averages") + 
  xlab("Year") + 
  ggtitle("Albert Pujols's Career mapped by Year") +
  geom_vline(xintercept = 2010, color = "Red")

a <- ggplot(Albert, aes(x = Year)) +
  geom_line(aes( y = HR), color = "blue") +
  geom_line(aes( y = SO), color = "red") + 
  geom_line(aes( y = BB), color = "purple") + 
  geom_line(aes( y = RBI), color = "Orange") +
  ylab("Totals By Category") + 
  xlab("Year") + 
  ggtitle("Albert Pujols's Career mapped by Year") +
  geom_vline(xintercept = 2010, color = "Black") + 
  annotate("Text", 2013.5, 130, label = "Decided to leave ALL his fans and \nleave St. Louis and head to the Angels. \nAll Stats Plummit")
 ggplotly(a)

Trout <- dat[grepl("Trout,M", dat$Player),]

b <- ggplot(Trout, aes(x = Year)) +
  geom_line(aes( y = AVG), color = "blue") +
  geom_line(aes( y = BABIP), color = "red") + 
  geom_line(aes( y = HRAB), color = "purple") + 
  geom_line(aes( y = RBIAB), color = "Orange") +
  ylab("Averages") + 
  xlab("Year") + 
  ggtitle("Mike Trout's Career mapped by Year") 

ggplot(Trout, aes(x = Year)) +
  geom_line(aes( y = HR), color = "blue") +
  geom_line(aes( y = SB), color = "red") + 
  geom_line(aes( y = BB), color = "purple") + 
  geom_line(aes( y = RBI), color = "Orange") +
  ylab("Totals By Category") + 
  xlab("Year") + 
  ggtitle("Mike Trout's Career mapped by Year") +
  geom_vline(xintercept = 2017, color = "Black")

```
Do Homerun hitters have higher Strike Out percentages?
```{r}
correlationMatrix2 <- select(dat, HRAB, SOAB)
cor(correlationMatrix2)

ggplot(dat, aes(x = HRAB, y = SOAB)) + geom_smooth(method='lm') + geom_point() + xlab("Home Runs Per at Bat") + ggtitle("Homerun Hitters and Strike Out Percentages") + ylab("Strike outs Per at Bat") + annotate("Text", .1225, .370, label = "      R = .402 \n R^2 = .162")


```

Graph rivalry teams by runs scored.
```{r}
RedSox <- dat[grepl("BOS", dat$Team),]
Yankees <- dat[grepl("NYY", dat$Team),]
Rivals <- full_join(RedSox, Yankees)

jColors = c("#BD3039", "#0C2340")
ggplot(Rivals, aes(x = HR, y = AVG)) + geom_boxplot() + facet_wrap(~Team) + xlab("Average Home Runs Hit in a Season") + ylab("Batting Average") + ggtitle("Rivaly Between the Boston Red Sox and New York Yankees") + aes(fill = Team) + scale_fill_manual(values = jColors)

```

Does Avg # of pitches in an at bat correlate with batting Avg? Does it correlate with HRs?

```{r}
##I CAN"T GET THE R STAT FOR THESE, SO WE WONT BE ABLE TO GET THE R^2

#This one proves the Null Hypothesis, that there is no Correlation
ggplot(dat, aes(x = AVG, y = NPAB)) + geom_smooth(method='lm') + geom_point() + xlab("Batting Average") + ggtitle("Relationship Between Average Number of Pitches in an at Bat to Batting Average") + ylab("Average Number of Pitches During an at Bat")## R_SQUARED


#This one has a slight positive correlation
ggplot(dat, aes(x = HRAB, y = NPAB)) + geom_smooth(method='lm') + geom_point() + xlab("Average Homeruns in an at Bat") + ggtitle("Relationship Between Average Number of Pitches in an at Bat to Batting Average") + ylab("Average Number of Pitches During an at Bat")## R_SQUARED
```

Distribution of BABIP with Buster Posey Highlighted

```{r}
Posey <- dat[grepl("Posey,B", dat$Player),]
MeanBABIPposey <- mean(Posey$BABIP)

ggplot(dat, aes(x = BABIP,)) + geom_histogram(color = "black", fill = "white") +
  geom_vline(xintercept = MeanBABIPposey, color = "Red") +
  annotate("Text", .38, 135, label = "Buster Posey's Average BABIP")

```

Which teams hit the most homeruns?
```{r}
dat7 <- dat


dat7%>%
  select(Team, HRAB) %>%
  mutate(HRAB = ifelse(HRAB == -9999, NA, HRAB)) %>%
  group_by(Team) %>%
  mutate(MeanHR = mean(HRAB, na.rm = T)) -> dat7

dat8 <- select(dat7, -HRAB)
dat8 <- distinct(dat8)  #ROUND TO 3 digits
dat8$MeanHR <- round(dat8$MeanHR, digits = 3)
head(dat8, 10)

dat8 <- dat8[order(dat8$MeanHR, decreasing = TRUE),] 

head(dat8, 33)

```

```{r}
library("Hmisc")
library(corrplot)
cordata <- select(dat, AVG, OBP, SLG, BABIP, OPS, HR, AO, SO, RBI, BB, XBH, NPPA, SB)
cordata <- cor(cordata, use = "complete.obs")
cordata.rcorr = rcorr(as.matrix(cordata))
cordata.rcorr
palette = colorRampPalette(c("blue", "yellow", "red")) (20)
corrplot(cordata, col = palette) #dot graph
heatmap(x = cordata, col = palette, symm = TRUE) #heatmap
#go back and change color scheme
```
Chloropleth map 
```{r}

Stadiums <- read_excel("StadiumLocations.xlsx")

View(Stadiums)

states <- map_data("state")
ggplot(states, aes(x=long, y=lat))+geom_polygon(aes(group=group))
ggplot(states, aes(x=long, y=lat))+geom_polygon(aes(group=group))+geom_point(aes(x=Longitude, y=Latitude), data=Stadiums %>% filter(Longitude <0 & Longitude> -130), size=2, color="red") 
#Change this to look better, put labels, put colors, 
#In the excel file, add averages for each team, make sure team names are able so we can display them


teamdat <- dat
teamdat%>%
  select(Team, AVG) %>%
  mutate(AVG = ifelse(AVG == -9999, NA, AVG)) %>%
  group_by(Team) %>%
  mutate(MEANAVG = mean(AVG, na.rm = T)) -> teamdat
ggplot(dat2, aes(x = AVG)) + geom_histogram(binwidth = .002,color="Black", fill="goldenrod2") + xlab("Batting Average") + ggtitle("Histogram of Batting Average by Position") + ylab("Count") + facet_wrap(~Pos) + geom_vline(aes(xintercept = MEANAVG))
```
